Moral Foundations of U.S. Political News Organizations

### 2022-03-03 manuscript_version/manuscript_metapsych.tex
update for tables why do they hate me
finally fixed the comma in the tables

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  english,
  man]{apa6}
\title{Moral Foundations of U.S. Political News Organizations}
\author{William E. Padfield\textsuperscript{1}, Erin M. Buchanan, Ph.D.\textsuperscript{2}, \& Kayla N. Jordan, Ph.D.\textsuperscript{2}}
\date{}

@@

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\HyOrg@maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother
\shorttitle{MORAL NEWS}
\keywords{politics, morality, psycholinguistics}
\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{csquotes}
\ifXeTeX
  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\else
  \usepackage[main=english]{babel}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
\fi
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

@@

\begin{table}[pt]

@@

\begin{tabular}{lcccccclccc}
\toprule
Source & $M_V$ & $SD_V$ & $N_{Article}$ & $N_{Words}$ & $M_T$ & $SD_T$ & $M_{Ty}$ & $SD_{Ty}$ & $M_{FK}$ & $SD_{FK}$\\
\midrule
NPR & 0.28 & 0.23 & 695 & 302977 & 435.94 & 642.63 & 191.96 & 192.28 & 14.00 & 3.93\\
New York Times & 0.30 & 0.13 & 406 & 452579 & 1,114.73 & 511.86 & 454.27 & 154.58 & 16.42 & 3.36\\
Breitbart & 0.29 & 0.18 & 1437 & 722022 & 502.45 & 347.90 & 243.35 & 120.75 & 18.57 & 7.89\\
Fox News & 0.29 & 0.17 & 503 & 296779 & 590.02 & 528.60 & 283.56 & 189.00 & 17.25 & 7.21\\
\bottomrule
\addlinespace
\end{tabular}

@@

\small{

\begin{tabular}{lcccccclccc}
\toprule
Source & $M_V$ & $SD_V$ & $N_{Article}$ & $N_{Words}$ & $M_T$ & $SD_T$ & $M_{Ty}$ & $SD_{Ty}$ & $M_{FK}$ & $SD_{FK}$\\
\midrule
Huffington Post & 0.27 & 0.12 & 552 & 359046 & 650.45 & 462.21 & 283.07 & 129.23 & 10.68 & 1.89\\
NPR & 0.22 & 0.25 & 366 & 108605 & 296.73 & 499.10 & 128.12 & 172.51 & 12.55 & 3.67\\
New York Times & 0.33 & 0.13 & 653 & 723569 & 1,108.07 & 570.31 & 461.26 & 174.20 & 9.52 & 1.79\\
Politico & 0.32 & 0.10 & 689 & 1069292 & 1,551.95 & 1,045.47 & 614.66 & 358.68 & 12.10 & 2.66\\
Slate & 0.29 & 0.12 & 272 & 229896 & 845.21 & 612.22 & 332.42 & 168.17 & 12.30 & 2.48\\
The Blaze & 0.25 & 0.13 & 277 & 128097 & 462.44 & 149.66 & 210.52 & 53.14 & 10.73 & 1.85\\
Breitbart & 0.30 & 0.13 & 757 & 375848 & 496.50 & 609.91 & 230.03 & 153.62 & 10.87 & 2.12\\
Fox News & 0.29 & 0.11 & 646 & 1304048 & 2,018.65 & 2,709.77 & 534.47 & 404.21 & 9.71 & 1.86\\
Sean Hannity & 0.31 & 0.11 & 27 & 5926 & 219.48 & 145.98 & 121.04 & 57.26 & 11.76 & 2.40\\
Rush Limbaugh & 0.38 & 0.13 & 172 & 267067 & 1,552.72 & 1,258.51 & 419.00 & 225.80 & 9.17 & 8.86\\
\bottomrule
\addlinespace
\end{tabular}

@@

\small{

\begin{tabular}{lcccccclccc}
\toprule
Source & $M_V$ & $SD_V$ & $N_{Article}$ & $N_{Words}$ & $M_T$ & $SD_T$ & $M_{Ty}$ & $SD_{Ty}$ & $M_{FK}$ & $SD_{FK}$\\
\midrule
Huffington Post & 0.30 & 0.13 & 432 & 242220 & 560.69 & 425.52 & 258.27 & 130.54 & 10.68 & 1.85\\
NPR & 0.23 & 0.22 & 434 & 151114 & 348.19 & 457.34 & 155.53 & 171.30 & 11.20 & 3.16\\
New York Times & 0.33 & 0.13 & 752 & 836811 & 1,112.78 & 532.72 & 472.95 & 177.57 & 10.10 & 1.74\\
Politico & 0.29 & 0.10 & 222 & 349499 & 1,574.32 & 875.63 & 638.32 & 313.34 & 11.31 & 1.26\\
Slate & 0.30 & 0.12 & 117 & 85254 & 728.67 & 464.34 & 301.71 & 138.74 & 11.87 & 2.53\\
The Blaze & 0.26 & 0.13 & 98 & 38983 & 397.79 & 102.10 & 185.66 & 41.93 & 10.70 & 1.85\\
Breitbart & 0.34 & 0.15 & 309 & 102886 & 332.96 & 238.70 & 168.31 & 81.59 & 10.71 & 2.05\\
Fox News & 0.35 & 0.13 & 1013 & 2799217 & 2,763.29 & 3,405.44 & 637.30 & 494.00 & 9.43 & 1.95\\
Sean Hannity & 0.29 & 0.14 & 63 & 10311 & 163.67 & 33.59 & 95.13 & 16.59 & 13.36 & 4.81\\
Rush Limbaugh & 0.37 & 0.12 & 78 & 152630 & 1,956.79 & 1,544.14 & 482.79 & 252.79 & 9.77 & 7.64\\
\bottomrule
\addlinespace
\end{tabular}

@@

For news articles about the partial U.S. Federal Government Shutdown of 2018-2019, there were no significant or practical effects of political lean for the moral foundations. A small-to-medium effect size was observed for \emph{authority/respect}. The effect was in the predicted direction as well, as conservative sources tended to offer more positive endorsements of the foundation than liberal sources. This effect was similar to Experiment 1 in which the largest effect size was observed for \emph{authority/respect}. As noted before, the effect found in Experiment 1 was in the opposite direction as what was hypothesized. Thus, it is difficult to draw comparisons between the two studies despite the similar pattern for effect size. Owing to a lack of similar effects for either Experiment 1 or the Kavanaugh topic, there is doubt as to whether or not a practical or generalized effect exists for \emph{authority/respect}.

\begin{table}[pt]

@@

\begin{table}[pt]

@@

@@

\leavevmode\vadjust pre{\hypertarget{ref-Clifford2013}{}}%
Clifford, S., \& Jerit, J. (2013). How words do the work of politics: Moral foundations theory and the debate over stem cell research. \emph{The Journal of Politics}, \emph{75}(3), 659--671.

@@

\leavevmode\vadjust pre{\hypertarget{ref-Frimer2020}{}}%
Frimer, J. A. (2020). Do liberals and conservatives use different moral languages? Two replications and six extensions of graham, haidt, and nosek's (2009) moral text analysis. \emph{Journal of Research in Personality}, \emph{84}, 103906.

\leavevmode\vadjust pre{\hypertarget{ref-Frimer2015}{}}%
Frimer, J. A., Tell, C. E., \& Haidt, J. (2015). Liberals condemn sacrilege too: The harmless desecration of cerro torre. \emph{Social Psychological and Personality Science}, \emph{6}(8), 878--886.

\leavevmode\vadjust pre{\hypertarget{ref-Garten2016}{}}%
Garten, J., Boghrati, R., Hoover, J., Johnson, K. M., \& Dehghani, M. (2016). Morality between the lines: Detecting moral sentiment in text. In \emph{Proceedings of IJCAI 2016 workshop on computational modeling of attitudes}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-Gray2015}{}}%
Gray, K., \& Keeney, J. E. (2015). Disconfirming moral foundations theory on its own terms: Reply to graham (2015). \emph{Social Psychological and Personality Science}, \emph{6}(8), 874--877.

\leavevmode\vadjust pre{\hypertarget{ref-Gray2014}{}}%
Gray, K., Schein, C., \& Ward, A. F. (2014). The myth of harmless wrongs in moral cognition: Automatic dyadic completion from sin to suffering. \emph{Journal of Experimental Psychology: General}, \emph{143}(4), 1600.

@@

\leavevmode\vadjust pre{\hypertarget{ref-Hopp2021}{}}%
Hopp, F. R., Fisher, J. T., Cornell, D., Huskey, R., \& Weber, R. (2021). The extended moral foundations dictionary (eMFD): Development and applications of a crowd-sourced approach to extracting moral intuitions from text. \emph{Behavior Research Methods}, \emph{53}(1), 232--246.

@@

\leavevmode\vadjust pre{\hypertarget{ref-Kincaid1975}{}}%
Kincaid, J. P., Fishburne, Jr., Robert P., R., Richard L., C., \& Brad S. (1975). {Derivation of New Readability Formulas (Automated Readability Index, Fog Count and Flesch Reading Ease Formula) for Navy Enlisted Personnel}. \url{https://doi.org/10.21236/ADA006655}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Sagi2014}{}}%
Sagi, E., \& Dehghani, M. (2014). Measuring moral rhetoric in text. \emph{Social Science Computer Review}, \emph{32}(2), 132--144.

\leavevmode\vadjust pre{\hypertarget{ref-Schein2015}{}}%
Schein, C., \& Gray, K. (2015). The unifying moral dyad: Liberals and conservatives share the same harm-based moral template. \emph{Personality and Social Psychology Bulletin}, \emph{41}(8), 1147--1163.

@@

\leavevmode\vadjust pre{\hypertarget{ref-Sterling2018}{}}%
Sterling, J., \& Jost, J. T. (2018). Moral discourse in the twitterverse: Effects of ideology and political sophistication on language use among US citizens and members of congress. \emph{Journal of Language and Politics}, \emph{17}(2), 195--221.

\leavevmode\vadjust pre{\hypertarget{ref-Suhler2011}{}}%
Suhler, C. L., Churchland, P., \& Joseph, C. (2011). {Can Innate , Modular {`` Foundations ''} Explain Morality ? Challenges for Haidt's Moral Foundations Theory}. \emph{Journal of Cognitive Neuroscience}, \emph{23}(9), 2103--2116.

@@

\leavevmode\vadjust pre{\hypertarget{ref-Wickham2016}{}}%
Wickham, H. (2016). {Package ` rvest '}. Retrieved from \url{https://cran.r-project.org/package=rvest}

@@

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  english,
  man,floatsintext]{apa6}
\title{Moral Foundations of U.S. Political News Organizations}
\author{William E. Padfield\textsuperscript{1}, Erin M. Buchanan, Ph.D.\textsuperscript{2}, \& Kayla N. Jordan, Ph.D.\textsuperscript{2}}
\date{}

@@

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\HyOrg@maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother
\shorttitle{MORAL NEWS}
\keywords{politics, morality, psycholinguistics}
\usepackage{csquotes}
\ifXeTeX
  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\else
  \usepackage[main=english]{babel}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
\fi
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

@@

\begin{table}[h]

@@

\begin{tabular}{lcccccccccc}
\toprule
Source & $M_V$ & $SD_V$ & $N_{Article}$ & $N_{Words}$ & $M_T$ & $SD_T$ & $M_{Ty}$ & $SD_{Ty}$ & $M_{FK}$ & $SD_{FK}$\\
\midrule
NPR & 0.28 & 0.23 & 695 & 302977 & 435.94 & 642.63 & 191.96 & 192.28 & 14.00 & 3.93\\
New York Times & 0.30 & 0.13 & 406 & 452579 & 1114.73 & 511.86 & 454.27 & 154.58 & 16.42 & 3.36\\
Breitbart & 0.29 & 0.18 & 1437 & 722022 & 502.45 & 347.90 & 243.35 & 120.75 & 18.57 & 7.89\\
Fox News & 0.29 & 0.17 & 503 & 296779 & 590.02 & 528.60 & 283.56 & 189.00 & 17.25 & 7.21\\
\bottomrule
\addlinespace
\end{tabular}

@@

\begin{tabular}{lcccccccccc}
\toprule
Source & $M_V$ & $SD_V$ & $N_{Article}$ & $N_{Words}$ & $M_T$ & $SD_T$ & $M_{Ty}$ & $SD_{Ty}$ & $M_{FK}$ & $SD_{FK}$\\
\midrule
Huffington Post & 0.27 & 0.12 & 552 & 359046 & 650.45 & 462.21 & 283.07 & 129.23 & 10.68 & 1.89\\
NPR & 0.22 & 0.25 & 366 & 108605 & 296.73 & 499.10 & 128.12 & 172.51 & 12.55 & 3.67\\
New York Times & 0.33 & 0.13 & 653 & 723569 & 1108.07 & 570.31 & 461.26 & 174.20 & 9.52 & 1.79\\
Politico & 0.32 & 0.10 & 689 & 1069292 & 1551.95 & 1045.47 & 614.66 & 358.68 & 12.10 & 2.66\\
Slate & 0.29 & 0.12 & 272 & 229896 & 845.21 & 612.22 & 332.42 & 168.17 & 12.30 & 2.48\\
The Blaze & 0.25 & 0.13 & 277 & 128097 & 462.44 & 149.66 & 210.52 & 53.14 & 10.73 & 1.85\\
Breitbart & 0.30 & 0.13 & 757 & 375848 & 496.50 & 609.91 & 230.03 & 153.62 & 10.87 & 2.12\\
Fox News & 0.29 & 0.11 & 646 & 1304048 & 2018.65 & 2709.77 & 534.47 & 404.21 & 9.71 & 1.86\\
Sean Hannity & 0.31 & 0.11 & 27 & 5926 & 219.48 & 145.98 & 121.04 & 57.26 & 11.76 & 2.40\\
Rush Limbaugh & 0.38 & 0.13 & 172 & 267067 & 1552.72 & 1258.51 & 419.00 & 225.80 & 9.17 & 8.86\\
\bottomrule
\addlinespace
\end{tabular}

@@

\begin{tabular}{lcccccccccc}
\toprule
Source & $M_V$ & $SD_V$ & $N_{Article}$ & $N_{Words}$ & $M_T$ & $SD_T$ & $M_{Ty}$ & $SD_{Ty}$ & $M_{FK}$ & $SD_{FK}$\\
\midrule
Huffington Post & 0.30 & 0.13 & 432 & 242220 & 560.69 & 425.52 & 258.27 & 130.54 & 10.68 & 1.85\\
NPR & 0.23 & 0.22 & 434 & 151114 & 348.19 & 457.34 & 155.53 & 171.30 & 11.20 & 3.16\\
New York Times & 0.33 & 0.13 & 752 & 836811 & 1112.78 & 532.72 & 472.95 & 177.57 & 10.10 & 1.74\\
Politico & 0.29 & 0.10 & 222 & 349499 & 1574.32 & 875.63 & 638.32 & 313.34 & 11.31 & 1.26\\
Slate & 0.30 & 0.12 & 117 & 85254 & 728.67 & 464.34 & 301.71 & 138.74 & 11.87 & 2.53\\
The Blaze & 0.26 & 0.13 & 98 & 38983 & 397.79 & 102.10 & 185.66 & 41.93 & 10.70 & 1.85\\
Breitbart & 0.34 & 0.15 & 309 & 102886 & 332.96 & 238.70 & 168.31 & 81.59 & 10.71 & 2.05\\
Fox News & 0.35 & 0.13 & 1013 & 2799217 & 2763.29 & 3405.44 & 637.30 & 494.00 & 9.43 & 1.95\\
Sean Hannity & 0.29 & 0.14 & 63 & 10311 & 163.67 & 33.59 & 95.13 & 16.59 & 13.36 & 4.81\\
Rush Limbaugh & 0.37 & 0.12 & 78 & 152630 & 1956.79 & 1544.14 & 482.79 & 252.79 & 9.77 & 7.64\\
\bottomrule
\addlinespace
\end{tabular}

@@

\begin{table}[h]

@@

For news articles about the partial U.S. Federal Government Shutdown of 2018-2019, there were no significant or practical effects of political lean for the moral foundations (see Table \ref{tab:exp2-tablegs}). A small-to-medium effect size was observed for \emph{authority/respect}. The effect was in the predicted direction as well, as conservative sources tended to offer more positive endorsements of the foundation than liberal sources. This effect was similar to Experiment 1 in which the largest effect size was observed for \emph{authority/respect}. As noted before, the effect found in Experiment 1 was in the opposite direction as what was hypothesized. Thus, it is difficult to draw comparisons between the two studies despite the similar pattern for effect size. Owing to a lack of similar effects for either Experiment 1 or the Kavanaugh topic, there is doubt as to whether or not a practical or generalized effect exists for \emph{authority/respect}.

\begin{table}[h]

@@

\newpage

\hypertarget{references}{%
\section{References}\label{references}}

\begingroup

@@

\leavevmode\vadjust pre{\hypertarget{ref-Clifford2013}{}}%
Clifford, S., \& Jerit, J. (2013). How words do the work of politics: Moral foundations theory and the debate over stem cell research. \emph{The Journal of Politics}, \emph{75}(3), 659--671. \url{https://doi.org/10.1017/S0022381613000492}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Frimer2020}{}}%
Frimer, J. A. (2020). Do liberals and conservatives use different moral languages? Two replications and six extensions of {G}raham, {H}aidt, and {N}osek's (2009) moral text analysis. \emph{Journal of Research in Personality}, \emph{84}, 103906. \url{https://doi.org/10.1016/j.jrp.2019.103906}

\leavevmode\vadjust pre{\hypertarget{ref-Frimer2015}{}}%
Frimer, J. A., Tell, C. E., \& Haidt, J. (2015). Liberals condemn sacrilege too: The harmless desecration of {C}erro {T}orre. \emph{Social Psychological and Personality Science}, \emph{6}(8), 878--886. \url{https://doi.org/10.1177/1948550615597974}

\leavevmode\vadjust pre{\hypertarget{ref-Garten2016}{}}%
Garten, J., Boghrati, R., Hoover, J., Johnson, K. M., \& Dehghani, M. (2016). Morality between the lines: Detecting moral sentiment in text. In \emph{{Proceedings of IJCAI 2016 workshop on Computational Modeling of Attitudes}}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-Gray2015}{}}%
Gray, K., \& Keeney, J. E. (2015). Disconfirming moral foundations theory on its own terms: Reply to {G}raham (2015). \emph{Social Psychological and Personality Science}, \emph{6}(8), 874--877. \url{https://doi.org/10.1177/1948550615592243}

\leavevmode\vadjust pre{\hypertarget{ref-Gray2014}{}}%
Gray, K., Schein, C., \& Ward, A. F. (2014). The myth of harmless wrongs in moral cognition: Automatic dyadic completion from sin to suffering. \emph{Journal of Experimental Psychology: General}, \emph{143}(4), 1600--1615. \url{https://doi.org/10.1037/a0036149}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Hopp2021}{}}%
Hopp, F. R., Fisher, J. T., Cornell, D., Huskey, R., \& Weber, R. (2021). The extended moral foundations dictionary (eMFD): Development and applications of a crowd-sourced approach to extracting moral intuitions from text. \emph{Behavior Research Methods}, \emph{53}(1), 232--246. \url{https://doi.org/10.3758/s13428-020-01433-0}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Kincaid1975}{}}%
Kincaid, J. P., Fishburne, Jr., Robert P., R., Richard L., C., \& Brad S. (1975). Derivation of new readability formulas (automated readability index, fog count and {F}lesch reading ease formula) for {N}avy enlisted personnel. \url{https://doi.org/10.21236/ADA006655}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Sagi2014}{}}%
Sagi, E., \& Dehghani, M. (2014). Measuring moral rhetoric in text. \emph{Social Science Computer Review}, \emph{32}(2), 132--144. \url{https://doi.org/10.1177/0894439313506837}

\leavevmode\vadjust pre{\hypertarget{ref-Schein2015}{}}%
Schein, C., \& Gray, K. (2015). The unifying moral dyad: Liberals and conservatives share the same harm-based moral template. \emph{Personality and Social Psychology Bulletin}, \emph{41}(8), 1147--1163. \url{https://doi.org/10.1177/0146167215591501}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Sterling2018}{}}%
Sterling, J., \& Jost, J. T. (2018). Moral discourse in the twitterverse: Effects of ideology and political sophistication on language use among US citizens and members of congress. \emph{Journal of Language and Politics}, \emph{17}(2), 195--221. \url{https://doi.org/10.1075/jlp.17034.ste}

\leavevmode\vadjust pre{\hypertarget{ref-Suhler2011}{}}%
Suhler, C. L., Churchland, P., \& Joseph, C. (2011). Can innate , modular "foundations" explain morality? Challenges for {H}aidt's {M}oral {F}oundations {T}heory. \emph{Journal of Cognitive Neuroscience}, \emph{23}(9), 2103--2116. \url{https://doi.org/10.1162/jocn.2011.21637}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Wickham2016}{}}%
Wickham, H. (2016). {Package `rvest'}. Retrieved from \url{https://cran.r-project.org/package=rvest}

@@

\endgroup
